---
title: "Airline_fatalitiesComparison"
author: "Caroline Oliver"
date: "4/3/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load in the dataset from csv file
### NOTE: Make sure to replace the path variable with your own path to the file
```{r}
library(readr)
airline_safety <- read_csv("/Users/carolineoliver/airline-safety.csv")
# airline_safety <- read_csv("path_to_csv_file_here")
# View(airline_safety)
```

### Get number of ASK based on trillions (like in the study)
```{r}
airline_safety$ASKperTrillion = airline_safety[, "avail_seat_km_per_week"]/1000000000000
#View(airline_safety)
```

### Get number of ASK based on billiosns to compare charts
```{r}
airline_safety$ASKperBillion = airline_safety[, "avail_seat_km_per_week"]/1000000000
#View(airline_safety)
```

# add columns to dataset for 85 to 99 (per trillion and per billion)
```{r}
airline_safety$Fatalities_85_99_trillion = airline_safety$fatalities_85_99 / airline_safety$ASKperTrillion
airline_safety$Fatalities_85_99_billion = airline_safety$fatalities_85_99 / airline_safety$ASKperBillion
```

### add columns to dataset for 00 to 14 (per trillion and per billion)
```{r}
airline_safety$Fatalities_00_14_trillion = airline_safety$fatalities_00_14 / airline_safety$ASKperTrillion
airline_safety$Fatalities_00_14_billion = airline_safety$fatalities_00_14 / airline_safety$ASKperBillion
#View(airline_safety)
```

### Make the plot for 'per trillion'
### the data for the columns needed is originally stored as a list 
### must be converted to numeric vector to be used in plot function
```{r}
tril_85_99 = as.numeric(as.character(unlist(airline_safety$Fatalities_85_99_trillion)))
tril_00_14 = as.numeric(as.character(unlist(airline_safety$Fatalities_00_14_trillion)))
plot(tril_85_99, tril_00_14, xlab="1985-99", ylab="2000-14", main = "Fatalities per Trillion")
abline(lm(tril_85_99 ~ tril_00_14))
```

### Make the plot for 'per billion'
### the data for the columns needed is originally stored as a list 
### must be converted to numeric vector to be used in plot function
```{r}
bil_85_99 = as.numeric(as.character(unlist(airline_safety$Fatalities_85_99_billion)))
bil_00_14 = as.numeric(as.character(unlist(airline_safety$Fatalities_00_14_billion)))
plot(bil_85_99, bil_00_14, xlab="1985-99", ylab="2000-14", main = "Fatalities per Billion")
abline(lm(bil_85_99 ~ bil_00_14))
```

## NOTES SO FAR ON GRAPHS ABOVE VS STUDY
### Per trillion does not yeild the same result axies as the study
### however, the graph points look the same pattern wise
### To attain the number of fatalities per trillion, I:
###     1) Divided total ASK given in first column by a trillion
###     2) Divided the number of fatalites by the numbers calculated in 1

### The same steps were followed to attain per billion numbers

### Both scatterplots show no correlation between the two year groups
### This is consistent with the study performed

### As stated above, all graphs appear to have the points pattern
### The only difference appears to be the axes

