---
title: "Predicting 2000-2014 Incidents using 1985-1999 data PLUS ADDED COLUMNS""
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load in the dataset from csv file
```{r}
library(randomForest)
library(tree)
library(e1071)
library(readr)

# NOTE: Make sure to replace the path variable with your own path to the file
airline_safety_new <- read.csv("/Users/carolineoliver/airline-safety-new.csv")
# REPLACE LINE ABOVE WITH YOUR PATH: airline_safety <- read.csv("path_to_csv_file_here")
```

#### Convert fatality numbers to 1 or 0
```{r}
i = 0
# Y_N stands for Yes_No
# Yes there were incidents = 1
# No there were no incidents = 0
incidentsY_N_00_14 = vector()

for (i in seq(nrow(airline_safety_new))){
  if(airline_safety_new$incidents_00_14[i] == 0){
    incidentsY_N_00_14[i] = 0
  }
  else{
    incidentsY_N_00_14[i] = 1
  }
}
```

#### Create new data frame for prediction
```{r}
incident_pred_df = airline_safety_new[ ,3:6]
incident_pred_df = cbind(incident_pred_df, airline_safety_new[ ,10:13])
View(incident_pred_df)

incident_pred_df$incidentsY_N_00_14 = incidentsY_N_00_14
colnames(incident_pred_df)[9] = "incidentsY_N_00_14"
View(incident_pred_df)
```

### Predict all airlines to have incidents - Error Rate
```{r}
table(incidentsY_N_00_14)

9 / 55
```

### Predict all airlines to NOT have incidents - Error Rate
```{r}
table(incidentsY_N_00_14)

46 / 55
```

### GLM Model
```{r}
glm_model = glm(incidentsY_N_00_14 ~ ., data=incident_pred_df)
summary(glm_model)

glm.pred<-predict(glm_model,incident_pred_df)
glm.pred
average_pred_value = sum(glm.pred)/55

# get 1 or 0 value for prediction
i = 0
predY_N = vector()

for (i in seq(nrow(airline_safety_new))){
  if(glm.pred[i] > average_pred_value){
    predY_N[i] = 1
  }
  else{
    predY_N[i] = 0
  }
}
```

#### GLM Confusion matrix
```{r}
table(predict=predY_N,truth=incidentsY_N_00_14)
```

#### GLM Error rate
```{r}
(24 + 0) / 24 + 0 + 9 + 22)
```

### Tree Classification Model
```{r fig2, fig.height = 8, fig.width = 10, fig.align = "center"}
incident.tree<-tree(incidentsY_N_00_14 ~ ., data=incident_pred_df)
summary(incident.tree)

plot(incident.tree)
text(incident.tree,pretty=0)

incident_tree.pred<-predict(incident.tree)
incident_tree.pred

i = 0
tree_predY_N = vector()

for (i in seq(nrow(airline_safety_new))){
  if(incident_tree.pred[i] > 0.5){
    tree_predY_N[i] = 1
  }
  else{
    tree_predY_N[i] = 0
  }
}
```

#### Tree Confusion Matrix
```{r}
table(tree_predY_N,incidentsY_N_00_14)
```

#### Tree Error Rate
```{r}
(4 + 4) / (4 + 4 + 28 + 19)
```

### Random Forest
```{r}
incident_pred_df$ incidentsY_N_00_14<-as.factor(fatal_pred_df$ fatalitiesY_N_00_14)
rf.incidents<-randomForest(incidentsY_N_00_14~.,data=fatal_pred_df,mtry=8,importance=T)
rf.incidents.pred<-predict(rf.incidents,incident_pred_df)
```

#### Random Forest Confusion Matrix
```{r}
table(rf.incidents.pred,incident_pred_df$incidentsY_N_00_14)
```

#### Random Forest Error Rate
```{r}
(0 + 0) / (23 + 32)
```
